Indexed Containers

### 2011-05-10 journal/paper.tex
approx

\noindent An elegant way to formalize and reason about inductive types
is to model them as the initial algebra of an endofunctor, we can
define the siganture functors cirresponding to each of the examples above:

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Conid{FNe}\;\in\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;\Conid{ℕ}\;\rightarrow\;\Conid{Set}{}\<[E]%
\\
\>[B]{}\Conid{FNe}\;\Conid{X}\;\Conid{Y}\;\Varid{n}\;\mathrel{=}\;\Conid{Fin}\;\Varid{n}\;\uplus\;(\Conid{X}\;\Varid{n}\;\uplus\;\Conid{Y}\;\Varid{n}){}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Conid{FNf}\;\in\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;\Conid{ℕ}\;\rightarrow\;\Conid{Set}{}\<[E]%
\\
\>[B]{}\Conid{FNf}\;\Conid{X}\;\Conid{Y}\;\Varid{n}\;\mathrel{=}\;(\Conid{Y}\;\ensuremath{\mbox{$\circ$}}\;\Varid{suc})\;\Varid{n}\;\uplus\;\Conid{X}\;\Varid{n}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

We have $\Pi$-types, denoted \ensuremath{(\Varid{a}\;\in\;\Conid{A})\;\rightarrow\;\Conid{B}\;\Varid{a}} and $\Sigma$-types, which we write as: \ensuremath{\Sigma(\!\;\Varid{a}\;\in\;\Conid{A}\;\!)\!\times\!\;\Conid{B}\;\Varid{a}}. In fact this is sugar for the record type:

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{data}\;\Conid{W}\;(\Conid{S}\;\in\;\Conid{Set})\;(\Conid{P}\;\in\;\Conid{S}\;\rightarrow\;\Conid{Set})\;\in\;\Conid{Set}\;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{sup}\;\in\;\Sigma(\!\;\Varid{s}\;\in\;\Conid{S}\;\!)\!\times\!\;(\Conid{P}\;\Varid{s}\;\rightarrow\;\Conid{W}\;\Conid{S}\;\Conid{P})\;\rightarrow\;\Conid{W}\;\Conid{S}\;\Conid{P}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{wrec}\;\in\;{}\<[9]%
\>[9]{}\{\mskip1.5mu \Conid{S}\;\in\;\Conid{Set}\mskip1.5mu\}\;\{\mskip1.5mu \Conid{P}\;\in\;\Conid{S}\;\rightarrow\;\Conid{Set}\mskip1.5mu\}\;(\Conid{Q}\;\in\;\Conid{W}\;\Conid{S}\;\Conid{P}\;\rightarrow\;\Conid{Set})\;{}\<[E]%
\\
\>[9]{}(\Varid{x}\;\in\;\Conid{W}\;\Conid{S}\;\Conid{P})\;{}\<[E]%
\\
\>[9]{}(\Varid{m}\;\in\;{}\<[15]%
\>[15]{}(\Varid{s}\;\in\;\Conid{S})\;(\Varid{f}\;\in\;\Conid{P}\;\Varid{s}\;\rightarrow\;\Conid{W}\;\Conid{S}\;\Conid{P})\;{}\<[E]%
\\
\>[15]{}(\Varid{h}\;\in\;(\Varid{p}\;\in\;\Conid{P}\;\Varid{s})\;\rightarrow\;\Conid{Q}\;(\Varid{f}\;\Varid{p}))\;{}\<[E]%
\\
\>[15]{}\rightarrow\;\Conid{Q}\;(\Varid{sup}\;(\Varid{s},\Varid{f})))\;{}\<[E]%
\\
\>[9]{}\rightarrow\;\Conid{Q}\;\Varid{x}{}\<[E]%
\\
\>[B]{}\Varid{wrec}\;\Conid{Q}\;(\Varid{sup}\;(\Varid{s},\Varid{f}))\;\Varid{m}\;\mathrel{=}\;\Varid{m}\;\Varid{s}\;\Varid{f}\;(\lambda\;\Varid{p}\;\rightarrow\;\Varid{wrec}\;\Conid{Q}\;(\Varid{f}\;\Varid{p})\;\Varid{m}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

\noindent
As a notational convenience, we will continue to define extra Agda data-types
in the rest of the paper, but in the end we will show how each of these can
be reduced to a theory that contains only \ensuremath{\Conid{W}}. For compactness, and
readability we will also define functions using Agda's pattern matching
syntax, rather than encoding them using \ensuremath{\Varid{wrec}}, it is an unstated lemma that
each of these definitions can be reduced to terms which only use \ensuremath{\Varid{wrec}}.

@@

We are going to use type theory versions of certain category theoretic
concepts For instance, we use ends \ensuremath{\Conid{End}} to capture natural transformations.
Given a bifunctor \ensuremath{\Conid{F}\;\in\;\Conid{Set}^{\Varid{op}}\;\rightarrow\;\Conid{Set}\;\rightarrow\;\Conid{Set}}, an element of \ensuremath{\prodd\;\Conid{X}\;.\;\Conid{F}\;\Conid{X}\;\Conid{X}} is
equivalent to an element of \ensuremath{\Varid{f}\;\in\;\{\mskip1.5mu \Conid{X}\;\in\;\Conid{Set}\mskip1.5mu\}\;\rightarrow\;\Conid{F}\;\Conid{X}\;\Conid{X}}, along with a proof:

@@

In this setting, the Yoneda lemma can be stated as follows, for any functor \ensuremath{\Conid{F}\;\in\;\Conid{Set}\;\rightarrow\;\Conid{Set}}:

@@

Finally, for readability we will elide certain artifacts in Agda's syntax,
for instance the quantification of implicit arguments when their types can be
inferred from the context. The reader should be reassured that the paper is a
literate agda file, available from the final author's webpage.

\todo{Define Func.}

@@

Every container \ensuremath{\Conid{S}\;\lhd\;\Conid{P}}  gives rise to a functor:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
\column{28}{@{}>{\hspre}l<{\hspost}@{}}%
\column{69}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{⟦\char95 ⟧}\;\in\;\Conid{Cont}\;\rightarrow\;\Conid{Func}{}\<[E]%
\\
\>[B]{}\Varid{⟦}\;\Conid{S}\;\lhd\;\Conid{P}\;\Varid{⟧}\;\mathrel{=}\;\Keyword{record}\;{}\<[21]%
\>[21]{}\{\mskip1.5mu \Varid{obj}\;{}\<[28]%
\>[28]{}\mathrel{=}\;\lambda\;\Conid{A}\;\rightarrow\;\Sigma(\!\;\Varid{s}\;\in\;\Conid{S}\;\!)\!\times\!\;(\Conid{P}\;\Varid{s}\;\rightarrow\;\Conid{A}){}\<[E]%
\\
\>[21]{};\Varid{mor}\;{}\<[28]%
\>[28]{}\mathrel{=}\;\lambda\;\Varid{m}\;\rightarrow\;\lambda (\!\;\Varid{s}\;\!,\!\;\Varid{f}\;\!\!)\;\rightarrow\;(\Varid{s},\Varid{m}\;\ensuremath{\mbox{$\circ$}}\;\Varid{f})\;{}\<[69]%
\>[69]{}\!\;\!{}\<[E]%
\\
\>[21]{}\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
\column{12}{@{}>{\hspre}c<{\hspost}@{}}%
\column{12E}{@{}l@{}}%
\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Sigma^{\text{\tiny F}}\;\in\;\forall{}\;\{\mskip1.5mu \Conid{J}\;\Conid{I}\;\Conid{K}\mskip1.5mu\}\;\rightarrow\;(\Conid{J}\;\rightarrow\;\Conid{K})\;\rightarrow\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{J}\;\rightarrow\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{K}{}\<[E]%
\\
\>[B]{}\Sigma^{\text{\tiny F}}\;\{\mskip1.5mu \Conid{J}\mskip1.5mu\}\;\Varid{f}\;\Conid{F}\;\Varid{k}\;\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{4}{}\<[4]%
\>[4]{}\Keyword{record}\;{}\<[12]%
\>[12]{}\{\mskip1.5mu {}\<[12E]%
\>[15]{}\Varid{obj}\;{}\<[20]%
\>[20]{}\mathrel{=}\;{}\<[23]%
\>[23]{}\lambda\;\Conid{A}\;\rightarrow\;\Sigma(\!\;\Varid{j}\;\in\;\Conid{J}\;\!)\!\times\!\;(\Varid{f}\;\Varid{j}\;\equiv\;\Varid{k}\;\times\;\;\Conid{F}\;\Conid{A}\;\Varid{j}){}\<[E]%
\\
\>[12]{};{}\<[12E]%
\>[15]{}\Varid{mor}\;{}\<[20]%
\>[20]{}\mathrel{=}\;{}\<[23]%
\>[23]{}\lambda\;\Varid{m}\;\rightarrow\;\lambda (\!\;\Varid{j}\;\!,\!\;\Varid{p}\;\!,\!\;\Varid{x}\;\!\!)\;\rightarrow\;(\Varid{j},\Varid{p},\;\Conid{F}\;\Varid{m}\;\Varid{j}\;\Varid{x})\;\!\;\!\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

\noindent
A parametrized \ensuremath{\Conid{F}}-algebra for \ensuremath{\Conid{F}\;\in\;\Conid{IFunc}^{\star}\;(\Conid{I}\;\uplus\;\Conid{J})\;\Conid{I}} is a pair of an 
indexed-functor \ensuremath{\Conid{G}\;\in\;\Conid{IFunc}\;\Conid{J}\;\Conid{I}} and a natural transformation
\ensuremath{\alpha\;\in\;\Conid{F}\;[\mskip1.5mu \;\Conid{G}\;\mskip1.5mu]^{\text{\tiny{F}}^{\star}}\;\Rightarrow^{\text{\tiny F}^{\star}}\;\Conid{G}}. A morphism between two such algebras 
\ensuremath{(\Conid{G},\alpha)} and \ensuremath{(\Conid{H},\beta)} is a natural transformation \ensuremath{\gamma\;\in\;\Conid{G}\;\Rightarrow^{\text{\tiny F}^{\star}}\;\Conid{H}} 
such that the follow diagram commutes:

@@

\noindent
As you might expect, a parametrized initial algebra for \ensuremath{\Conid{F}}, if it is exists, 
will be the initial object in the category of parametrized \ensuremath{\Conid{F}}-algebras.

The fact that the parametrized initial algebra construction can be iterated, 
means that we can define nested and mutual families of data-types, such as the
tuple of neutral and normal \ensuremath{\lambda}-terms outlined in the introduction.

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{11}{@{}>{\hspre}c<{\hspost}@{}}%
\column{11E}{@{}l@{}}%
\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
\column{19}{@{}>{\hspre}l<{\hspost}@{}}%
\column{26}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{⟦\char95 ⟧}\;\in\;\forall{}\;\{\mskip1.5mu \Conid{I}\mskip1.5mu\}\;\rightarrow\;\Conid{ICont}\;\Conid{I}\;\rightarrow\;\Conid{IFunc}\;\Conid{I}{}\<[E]%
\\
\>[B]{}\Varid{⟦\char95 ⟧}\;\{\mskip1.5mu \Conid{I}\mskip1.5mu\}\;(\Conid{S}\;\lhd\;\Conid{P})\;\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Keyword{record}\;{}\<[11]%
\>[11]{}\{\mskip1.5mu {}\<[11E]%
\>[14]{}\Varid{obj}\;{}\<[19]%
\>[19]{}\mathrel{=}\;\lambda\;\Conid{A}\;{}\<[26]%
\>[26]{}\rightarrow\;\Sigma(\!\;\Varid{s}\;\in\;\Conid{S}\;\!)\!\times\!\;(\Conid{P}\;\Varid{s}\;\rightarrow^{\star}\;\Conid{A}){}\<[E]%
\\
\>[11]{};{}\<[11E]%
\>[14]{}\Varid{mor}\;{}\<[19]%
\>[19]{}\mathrel{=}\;\lambda\;\Varid{m}\;{}\<[26]%
\>[26]{}\rightarrow\;\lambda (\!\;\Varid{s}\;\!,\!\;\Varid{f}\;\!\!)\;\rightarrow\;(\Varid{s},\Varid{m}\;\circ^{\star}\;\Varid{f})\;\!\;\!\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

\begin{align*}
                & \ensuremath{\Varid{⟦}\;\Conid{S}\;\lhd\;\Conid{P}\;\Varid{⟧}\;\Rightarrow^{\text{\tiny F}}\;\Conid{F}} & \\
  \equiv  \;    & \ensuremath{\prodd\;\Conid{X}\;.\;\Sigma(\!\;\Varid{s}\;\in\;\Conid{S}\;(\Conid{P}\;\Varid{s}\;\rightarrow^{\star}\;\Conid{X})\;\rightarrow\;\Conid{F}\;\Conid{X}} & \{\mbox{by definition}\} \\
  \equiv  \;    & \ensuremath{\prodd\;\Conid{X}\;.\;(\Varid{s}\;\in\;\Varid{s})\;\rightarrow\;(\Conid{P}\;\Varid{s}\;\rightarrow^{\star}\;\Conid{X})\;\rightarrow\;\Conid{F}\;\Conid{X}} & \{\mbox{currying}\} \\
  \cong   \;    & \ensuremath{(\Varid{s}\;\in\;\Conid{S})\;\rightarrow\;\prodd\;\Conid{X}\;.\;(\Conid{P}\;\Varid{s}\;\rightarrow^{\star}\;\Conid{X})\;\rightarrow\;\Conid{F}\;\Conid{X}} & \{\mbox{commuting end and pi} \} \\
  \cong   \;    & \ensuremath{(\Varid{s}\;\in\;\Conid{S})\;\rightarrow\;\Conid{F}\;(\Conid{P}\;\Varid{s})} & \{\mbox{Yoneda}\} \\
\end{align*}

>>>>

\noindent An elegant way to formalize and reason about inductive types
is to model them as the initial algebra of an endofunctor, we can
define the siganture functors corresponding to each of the examples above:

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Conid{F}_{\Conid{NeLam}}\;\in\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;\Conid{ℕ}\;\rightarrow\;\Conid{Set}{}\<[E]%
\\
\>[B]{}\Conid{F}_{\Conid{NeLam}}\;\Conid{X}\;\Conid{Y}\;\Varid{n}\;\mathrel{=}\;\Conid{Fin}\;\Varid{n}\;\uplus\;(\Conid{X}\;\Varid{n}\;\uplus\;\Conid{Y}\;\Varid{n}){}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Conid{F}_{\Conid{NfLam}}\;\in\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;(\Conid{ℕ}\;\rightarrow\;\Conid{Set})\;\rightarrow\;\Conid{ℕ}\;\rightarrow\;\Conid{Set}{}\<[E]%
\\
\>[B]{}\Conid{F}_{\Conid{NfLam}}\;\Conid{X}\;\Conid{Y}\;\Varid{n}\;\mathrel{=}\;(\Conid{Y}\;\ensuremath{\mbox{$\circ$}}\;\Varid{suc})\;\Varid{n}\;\uplus\;\Conid{X}\;\Varid{n}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

We have $\Pi$-types, denoted \ensuremath{(\Varid{a}\;\in\;\Conid{A})\;\rightarrow\;\Conid{B}\;\Varid{a}} and $\Sigma$-types, which we write as: \ensuremath{(\!\;\Varid{a}\;\in\;\Conid{A}\;\!)\times\;\Conid{B}\;\Varid{a}}. In fact this is sugar for the record type:

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{data}\;\Conid{W}\;(\Conid{S}\;\in\;\Conid{Set})\;(\Conid{P}\;\in\;\Conid{S}\;\rightarrow\;\Conid{Set})\;\in\;\Conid{Set}\;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{sup}\;\in\;(\!\;\Varid{s}\;\in\;\Conid{S}\;\!)\times\;(\Conid{P}\;\Varid{s}\;\rightarrow\;\Conid{W}\;\Conid{S}\;\Conid{P})\;\rightarrow\;\Conid{W}\;\Conid{S}\;\Conid{P}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{wrec}\;\in\;{}\<[9]%
\>[9]{}\{\mskip1.5mu \Conid{S}\;\in\;\Conid{Set}\mskip1.5mu\}\;\{\mskip1.5mu \Conid{P}\;\in\;\Conid{S}\;\rightarrow\;\Conid{Set}\mskip1.5mu\}\;(\Conid{Q}\;\in\;\Conid{W}\;\Conid{S}\;\Conid{P}\;\rightarrow\;\Conid{Set})\;{}\<[E]%
\\
\>[9]{}(\Varid{x}\;\in\;\Conid{W}\;\Conid{S}\;\Conid{P})\;{}\<[E]%
\\
\>[9]{}(\Varid{m}\;\in\;{}\<[15]%
\>[15]{}(\Varid{s}\;\in\;\Conid{S})\;(\Varid{f}\;\in\;\Conid{P}\;\Varid{s}\;\rightarrow\;\Conid{W}\;\Conid{S}\;\Conid{P})\;{}\<[E]%
\\
\>[15]{}(\Varid{h}\;\in\;(\Varid{p}\;\in\;\Conid{P}\;\Varid{s})\;\rightarrow\;\Conid{Q}\;(\Varid{f}\;\Varid{p}))\;{}\<[E]%
\\
\>[15]{}\rightarrow\;\Conid{Q}\;(\Varid{sup}\;(\Varid{s},\Varid{f})))\;{}\<[E]%
\\
\>[9]{}\rightarrow\;\Conid{Q}\;\Varid{x}{}\<[E]%
\\
\>[B]{}\Varid{wrec}\;\Conid{Q}\;(\Varid{sup}\;(\Varid{s},\Varid{f}))\;\Varid{m}\;\mathrel{=}\;\Varid{m}\;\Varid{s}\;\Varid{f}\;(\lambda\;\Varid{p}\;\rightarrow\;\Varid{wrec}\;\Conid{Q}\;(\Varid{f}\;\Varid{p})\;\Varid{m}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

\noindent
As a notational convenience, we will continue to define extra Agda data-types
in the rest of the paper, but in the end we will show how each of these can
be reduced to a theory that contains only \ensuremath{\Conid{W}}. For compactness, and
readability we will also define functions using Agda's pattern matching
syntax, rather than encoding them using \ensuremath{\Varid{wrec}}, all of these definitions can 
be reduced to terms which only use \ensuremath{\Varid{wrec}}.

@@

We are going to use type theoretic versions of certain category theoretic
concepts. For example we represent functors by packing up their definitions in Agda's records. An endo functor on set, is given by:

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{8}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{record}\;\Conid{Func}\;\in\;\Conid{Set₁}\;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{obj}\;{}\<[8]%
\>[8]{}\in\;\Conid{Set}\;\rightarrow\;\Conid{Set}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{mor}\;{}\<[8]%
\>[8]{}\in\;\forall{}\;\{\mskip1.5mu \Conid{A}\;\Conid{B}\mskip1.5mu\}\;\rightarrow\;(\Conid{A}\;\rightarrow\;\Conid{B})\;\rightarrow\;\Varid{obj}\;\Conid{A}\;\rightarrow\;\Varid{obj}\;\Conid{B}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

It would also be possible to pack up the functor laws as extra fields in these 
records.

We use ends \ensuremath{\Conid{End}} to capture natural transformations.
Given a bifunctor \ensuremath{\Conid{F}\;\in\;\Conid{Set}^{\Varid{op}}\;\rightarrow\;\Conid{Set}\;\rightarrow\;\Conid{Set}}, an element of \ensuremath{\prodd\;\Conid{X}\;.\;\Conid{F}\;\Conid{X}\;\Conid{X}} is
equivalent to an element of \ensuremath{\Varid{f}\;\in\;\{\mskip1.5mu \Conid{X}\;\in\;\Conid{Set}\mskip1.5mu\}\;\rightarrow\;\Conid{F}\;\Conid{X}\;\Conid{X}}, along with a proof:

@@

In this setting, the Yoneda lemma can be stated as follows, for any functor \ensuremath{\Conid{F}}:

@@

Finally, for readability we will elide certain artifacts in Agda's syntax,
for instance the quantification of implicit arguments when their types can be
inferred from the context. We will often leave out record projections 
from notions such as \ensuremath{\Conid{Func}}, allowing the functor to stand for both its action 
on object and morphism, just as would happen in the category theory 
literature. The reader should be reassured that the paper is a
literate agda file, available from the final author's webpage.

@@

Every container \ensuremath{\Conid{S}\;\lhd\;\Conid{P}}  gives rise to a functor:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
\column{28}{@{}>{\hspre}l<{\hspost}@{}}%
\column{69}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{⟦\char95 ⟧}\;\in\;\Conid{Cont}\;\rightarrow\;\Conid{Func}{}\<[E]%
\\
\>[B]{}\Varid{⟦}\;\Conid{S}\;\lhd\;\Conid{P}\;\Varid{⟧}\;\mathrel{=}\;\Keyword{record}\;{}\<[21]%
\>[21]{}\{\mskip1.5mu \Varid{obj}\;{}\<[28]%
\>[28]{}\mathrel{=}\;\lambda\;\Conid{A}\;\rightarrow\;(\!\;\Varid{s}\;\in\;\Conid{S}\;\!)\times\;(\Conid{P}\;\Varid{s}\;\rightarrow\;\Conid{A}){}\<[E]%
\\
\>[21]{};\Varid{mor}\;{}\<[28]%
\>[28]{}\mathrel{=}\;\lambda\;\Varid{m}\;\rightarrow\;\lambda (\!\;\Varid{s}\;\!,\!\;\Varid{f}\;\!\!)\;\rightarrow\;(\Varid{s},\Varid{m}\;\ensuremath{\mbox{$\circ$}}\;\Varid{f})\;{}\<[69]%
\>[69]{}\!\;\!{}\<[E]%
\\
\>[21]{}\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
\column{12}{@{}>{\hspre}c<{\hspost}@{}}%
\column{12E}{@{}l@{}}%
\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Sigma^{\text{\tiny F}}\;\in\;\forall{}\;\{\mskip1.5mu \Conid{J}\;\Conid{I}\;\Conid{K}\mskip1.5mu\}\;\rightarrow\;(\Conid{J}\;\rightarrow\;\Conid{K})\;\rightarrow\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{J}\;\rightarrow\;\Conid{IFunc}^{\star}\;\Conid{I}\;\Conid{K}{}\<[E]%
\\
\>[B]{}\Sigma^{\text{\tiny F}}\;\{\mskip1.5mu \Conid{J}\mskip1.5mu\}\;\Varid{f}\;\Conid{F}\;\Varid{k}\;\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{4}{}\<[4]%
\>[4]{}\Keyword{record}\;{}\<[12]%
\>[12]{}\{\mskip1.5mu {}\<[12E]%
\>[15]{}\Varid{obj}\;{}\<[20]%
\>[20]{}\mathrel{=}\;{}\<[23]%
\>[23]{}\lambda\;\Conid{A}\;\rightarrow\;(\!\;\Varid{j}\;\in\;\Conid{J}\;\!)\times\;(\Varid{f}\;\Varid{j}\;\equiv\;\Varid{k}\;\times\;\;\Conid{F}\;\Conid{A}\;\Varid{j}){}\<[E]%
\\
\>[12]{};{}\<[12E]%
\>[15]{}\Varid{mor}\;{}\<[20]%
\>[20]{}\mathrel{=}\;{}\<[23]%
\>[23]{}\lambda\;\Varid{m}\;\rightarrow\;\lambda (\!\;\Varid{j}\;\!,\!\;\Varid{p}\;\!,\!\;\Varid{x}\;\!\!)\;\rightarrow\;(\Varid{j},\Varid{p},\;\Conid{F}\;\Varid{m}\;\Varid{j}\;\Varid{x})\;\!\;\!\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

\noindent
A parametrised \ensuremath{\Conid{F}}-algebra for \ensuremath{\Conid{F}\;\in\;\Conid{IFunc}^{\star}\;(\Conid{I}\;\uplus\;\Conid{J})\;\Conid{I}} is a pair of an 
indexed-functor \ensuremath{\Conid{G}\;\in\;\Conid{IFunc}\;\Conid{J}\;\Conid{I}} and a natural transformation
\ensuremath{\alpha\;\in\;\Conid{F}\;[\mskip1.5mu \;\Conid{G}\;\mskip1.5mu]^{\text{\tiny{F}}^{\star}}\;\Rightarrow^{\text{\tiny F}^{\star}}\;\Conid{G}}. A morphism between two such algebras 
\ensuremath{(\Conid{G},\alpha)} and \ensuremath{(\Conid{H},\beta)} is a natural transformation \ensuremath{\gamma\;\in\;\Conid{G}\;\Rightarrow^{\text{\tiny F}^{\star}}\;\Conid{H}} 
such that the follow diagram commutes:

@@

\noindent
As you might expect, a parametrised initial algebra for \ensuremath{\Conid{F}}, if it is exists, 
will be the initial object in the category of parametrised \ensuremath{\Conid{F}}-algebras.

The fact that the parametrised initial algebra construction can be iterated, 
means that we can define nested and mutual families of data-types, such as the
tuple of neutral and normal \ensuremath{\lambda}-terms outlined in the introduction.

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{11}{@{}>{\hspre}c<{\hspost}@{}}%
\column{11E}{@{}l@{}}%
\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
\column{19}{@{}>{\hspre}l<{\hspost}@{}}%
\column{26}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{⟦\char95 ⟧}\;\in\;\forall{}\;\{\mskip1.5mu \Conid{I}\mskip1.5mu\}\;\rightarrow\;\Conid{ICont}\;\Conid{I}\;\rightarrow\;\Conid{IFunc}\;\Conid{I}{}\<[E]%
\\
\>[B]{}\Varid{⟦\char95 ⟧}\;\{\mskip1.5mu \Conid{I}\mskip1.5mu\}\;(\Conid{S}\;\lhd\;\Conid{P})\;\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Keyword{record}\;{}\<[11]%
\>[11]{}\{\mskip1.5mu {}\<[11E]%
\>[14]{}\Varid{obj}\;{}\<[19]%
\>[19]{}\mathrel{=}\;\lambda\;\Conid{A}\;{}\<[26]%
\>[26]{}\rightarrow\;(\!\;\Varid{s}\;\in\;\Conid{S}\;\!)\times\;(\Conid{P}\;\Varid{s}\;\rightarrow^{\star}\;\Conid{A}){}\<[E]%
\\
\>[11]{};{}\<[11E]%
\>[14]{}\Varid{mor}\;{}\<[19]%
\>[19]{}\mathrel{=}\;\lambda\;\Varid{m}\;{}\<[26]%
\>[26]{}\rightarrow\;\lambda (\!\;\Varid{s}\;\!,\!\;\Varid{f}\;\!\!)\;\rightarrow\;(\Varid{s},\Varid{m}\;\circ^{\star}\;\Varid{f})\;\!\;\!\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

\begin{align*}
                & \ensuremath{\Varid{⟦}\;\Conid{S}\;\lhd\;\Conid{P}\;\Varid{⟧}\;\Rightarrow^{\text{\tiny F}}\;\Conid{F}} & \\
  \equiv  \;    & \ensuremath{\prodd\;\Conid{X}\;.\;(\!\;\Varid{s}\;\in\;\Conid{S}\;(\Conid{P}\;\Varid{s}\;\rightarrow^{\star}\;\Conid{X})\;\rightarrow\;\Conid{F}\;\Conid{X}} & \{\mbox{by definition}\} \\
  \equiv  \;    & \ensuremath{\prodd\;\Conid{X}\;.\;(\Varid{s}\;\in\;\Varid{s})\;\rightarrow\;(\Conid{P}\;\Varid{s}\;\rightarrow^{\star}\;\Conid{X})\;\rightarrow\;\Conid{F}\;\Conid{X}} & \{\mbox{currying}\} \\
  \cong   \;    & \ensuremath{(\Varid{s}\;\in\;\Conid{S})\;\rightarrow\;\prodd\;\Conid{X}\;.\;(\Conid{P}\;\Varid{s}\;\rightarrow^{\star}\;\Conid{X})\;\rightarrow\;\Conid{F}\;\Conid{X}} & \{\mbox{commuting end and pi} \} \\
  \cong   \;    & \ensuremath{(\Varid{s}\;\in\;\Conid{S})\;\rightarrow\;\Conid{F}\;(\Conid{P}\;\Varid{s})} & \{\mbox{Yoneda}\} \\
\end{align*}
